Benchmarking on Rigetti architecture

### 2022-10-01 txt/manuscript.tex
preli

\subsubsection{Model specification}
The scenario assumes that Alice and Bob have access to an unknown measurement
device, a black box. Alice has hidden one of the two measurements, either $\SS$
or $\TT$, in the black box. Bob only has the information that the black box
performs one of these measurements. They want to know the upper bound on the
probability of correct discrimination between $\SS$ and $\TT$. They also want to
construct the optimal discrimination strategy to achieve the highest probability
of correct discrimination. Theoretical results on discrimination of von Neumann
measurements are characterized in~\cite{puchala2018strategies}.

In this work we will consider the problem of discrimination between von Neumann 
measurement in the computational basis and a von Neumann measurement in basis 
$U$, where $U$ is a unitary matrix. 
The scheme of discrimination can be divided into the following steps, which are 
illustrated in Fig.~\ref{fig:theoretical_scheme}.

@@

\begin{enumerate}
	\item Alice and Bob share an entangled state $\ket{\psi_{AB}}$ on systems 
	$A$ and $B$.
	\item Alice performs one of two known measurements (either $\PP_{\Id}$ or 
	$\PP_{U}$) on part $A$ of the input state  $\ket{\psi_{AB}}$.
	\item Alice performs the conditional binary measurement on the part 
	$B$ based on the output label $i$. 
	\item  Bob measures part $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision based on the received label $j$. If $j=0$, then he decides 
	that the the black box contains $\PP_U$. Otherwise, he decides that the 
	black box contains $\PP_{\Id}$. 
\end{enumerate}

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/theoretical_scheme}
	\caption{Theoretical discrimination scheme}
	\label{fig:theoretical_scheme}
\end{figure}

@@

Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:theoretical_scheme}). One of the components of 
this scheme is to perform the conditional binary measurement. 
Unfortunately, current NISQ devices do not allow for performing conditional 
binary measurements.
The only possible measurement performed on quantum architectures is the 
measurement in the computational basis. 
We  will present two possible methods providing a clever way to cross this 
obstacle. These will be two schemes of practical realization. 
The first method consists in using post-selection while the second method 
consists in applying a controlled unitary. These methods are described below.

\begin{scheme}(By using postsellection)

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{AB}}$ on systems 
$A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or 
$\Phi_{U}$,  on part $A$ of the input state  $\ket{\psi_{AB}}$.
\item Alice measures part $A$ of the input state  $\ket{\psi_{AB}}$ and 
receives an output label $i$.
\item  
From Holevo--Helstrom theorem, Bob performs the conditional binary 
measurement	$\PP_{V_j^\dagger}$ on part $B$.
\item They include the cases for which $i = j$.\todo{?} Otherwise, they reject 
them. 
\item Bob measures part $B$ of the input state  $\ket{\psi_{AB}}$ and makes a
decision based on received label $j$. If $j=0$, then he decides that the the
black box contains $\PP_U$. Otherwise, he decides that the black box contains
$\PP_{\Id}$.
\end{enumerate}

@@

PROBABILITY 
		\begin{equation}
		p = ...
		\end{equation}
		%\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{AB}}$ on 
systems $A$ and $B$..
%\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
%\rangle. $
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or
$\Phi_{U}$,  on part $A$ of the input state  $\ket{\psi_{AB}}$.
\item Alice and Bob implement controlled unitary $V_0 \oplus V_1$ obtained from 
Holevo--Helstrom theorem.	
\item Alice and Bob prepare the measurement $\Delta$ in computational basis on 
their systems.
\item Bob makes a decision based on received label $j$. If $j=0$, then he
decides that the the black box contains $\PP_U$. Otherwise, he decides that the
black box contains $\PP_{\Id}$.
\end{enumerate}

@@

PROBABILITY

\begin{equation}
p = \frac{|j=0 \wedge \Phi_{U} | + |j=1 \wedge \Phi_{\1}|}{\text{all 
		cases}}
\end{equation}

@@

Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
A unitary channel
$\Phi_{U}$ is defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$. A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix and hence we will use the notation $\PP_{U}$
for a von Neumann measurement with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i+1$-th column of the unitary matrix $U$. The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
seen as  a measure-and-prepare quantum channel as follows 
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$ It is worth noting that a given  von Neumann measurement  $\PP_{U}$ can be also rewritten as $\Delta 
\circ \Phi_{(UE)^\dagger}$, where $\Delta$ denotes the completely dephasing 
channel.

>>>>

\subsubsection{Model specification}
The scenario assumes that Alice and Bob have access to an unknown measurement
device, a black box. Alice has hidden one of the two measurements, either $\PP$
or $\mathcal{Q}$, in the black box. Bob only has the information that the black box
performs one of these measurements. They want to determine an upper bound on the
probability of correct discrimination between $\PP$ and $\mathcal{Q}$. Moreover, they also want to
construct an optimal discrimination strategy to achieve the highest probability. Theoretical results were characterized in~\cite{puchala2018strategies}.

In this work, we will consider the problem of discrimination between von Neumann 
measurements. One of them on the computational basis, the second one on basis 
$U$, where $U$ is a unitary matrix. 
The scheme of discrimination can be divided into the following steps, which are 
illustrated in Fig.~\ref{fig:theoretical_scheme}.

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/theoretical_scheme}
	\caption{Theoretical  scheme of discrimination  between von Neumann measurements $\PP_{U}$ and $\PP_\Id$. }
	\label{fig:theoretical_scheme}
\end{figure}

@@

\begin{enumerate}
	\item Alice and Bob share an entangled state $\ket{\psi_{AB}}$ on systems 
	$A$ and $B$.
	\item Alice performs one of two known measurements (either $\PP_{\Id}$ or 
	$\PP_{U}$) on part $A$ of the input state  $\ket{\psi_{AB}}$ and measures the system $A$.
	\item Alice performs the conditional binary measurement $\PP_{V_i}$ on the part 
	$B$ based on the output label $i$. 
	\item  Bob measures the system $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision based on the received label $j$. If $j=0$, then he decides 
	that the the black box contains $\PP_U$. Otherwise, he decides that the 
	black box contains $\PP_{\Id}$. 
\end{enumerate}

@@

Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:theoretical_scheme}). One of the components of 
this scheme is to perform a conditional binary measurement. 
Unfortunately, current NISQ devices do not allow for performing conditional 
binary measurements.
The only possible measurement performed on quantum architectures is the 
measurement in the computational basis. 
We  will present two possible methods providing a clever way to cross this 
obstacle. These will be two schemes of practical realization. 
The first method consists in using postselection while the second method 
consists in applying a controlled unitary. These methods are described below.

\begin{scheme}(Postselection)

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{AB}}$ on systems 
$A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or 
$\Phi_{U}$,  on part $A$ of the input state  $\ket{\psi_{AB}}$.
\item Alice measures part $A$ of the input state  $\ket{\psi_{AB}}$ and 
receives an output label $i$.
\item  
Based on Holevo--Helstrom theorem, Bob performs a conditional binary 
measurement	$\PP_{V_j^\dagger}$ on part $B$.
\item To calculate the probability of correct discrimination, they include just those cases for which $i = j$.
\item Bob measures part $B$ of the input state  $\ket{\psi_{AB}}$ and makes a
decision based on received label $j$. If $i=j=0$, then he decides that the 
black box contains $\PP_U$. Otherwise, he decides that the black box contains
$\PP_{\Id}$.
\end{enumerate}

@@

Due to the fact that the measurement result is nondeterministic, we are forced to consider the following circuits
in the postselection procedure
\begin{equation}
\begin{split}
(\Phi_{U}, \mathcal{P}_{V_k}, i, j) \\ 
(\Phi_{\Id}, \mathcal{P}_{V_k}, i, j)
\end{split}
\end{equation}
where $k=\{0,1\}, i=\{0,1\}$ and $ j=\{0,1\}$. 
The optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
		\begin{equation}
		p = \frac{\#(\Phi_U, \PP_{V_i},i,0) + \#(\Phi_\Id, \PP_{V_i},i,1)}{\#(\Phi_U, \PP_{V_i},i,j) + \#(\Phi_\Id, \PP_{V_i},i,j)}, 
		\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$.

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{AB}}$ on 
systems $A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or
$\Phi_{U}$,  on part $A$ of the input state  $\ket{\psi_{AB}}$.
\item Alice and Bob implement controlled unitary $V_0 \oplus V_1$ obtained from 
Holevo--Helstrom theorem.	
\item Alice and Bob prepare the measurement $\Delta$ in computational basis on 
their systems.
\item Bob makes a decision based on received label $j$. If $j=0$, then he
decides that the the black box contains $\PP_U$. Otherwise, he decides that the
black box contains $\PP_{\Id}$.
\end{enumerate}

@@

Here, we consider the following circuits
by using controlled unitary
\begin{equation}
\begin{split}
(\Phi_{U}, V_0 \oplus V_1, i, j) \\ 
(\Phi_{\Id},  V_0 \oplus V_1, i, j)
\end{split}
\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$. 
Then, the optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
\begin{equation}
p = \frac{\#(\Phi_U,V_0 \oplus V_1,i,0) + \#(\Phi_\Id,V_0 \oplus V_1,i,1) }{\#(\Phi_U,V_0 \oplus V_1,i,j) + \#(\Phi_\Id,V_0 \oplus V_1,i,j)}
\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$. 
\end{scheme}

@@

Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
In this work we will consider a special class of quantum channels, called unitary channels. A unitary channel
$\Phi_{U}$ is defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$. A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix and hence we will use the notation $\PP_{U}$
for a von Neumann measurement with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i+1$-th column of the unitary matrix $U$. The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
seen as  a measure-and-prepare quantum channel as follows 
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$
